---
title: "Tesla Stock Price Prediction"
author: "ERICK@Guru"
date: "2024-04"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Tesla Stock Price Prediction

In this project, **Exploratory Data Analysis** is carried out on TESLA Stock data obtained from Yahoo Finande:

Here, EDA is performed with different data analysis technologies viz. Python, R, PowerBI, and Microsoft Excel. Furthermore, python keras package is used to predict the stock prices for the month of January and also predicted percentage increase or decrease in the stock prices.

```{r warning=FALSE, message=FALSE}
#import libraries
library(tidyverse)
library(ggplot2)
library(zoo)

```

```{r warning=FALSE, message=FALSE}
#import dataset
TSLA <- read_csv("TSLA.csv")
sum(duplicated(TSLA$Date))
colSums(is.na(TSLA))
glimpse(TSLA)
```

```{r}
#manipulate date variable
TSLA$Date <- ymd(TSLA$Date)
```

### Basic trend Analysis using Closing prices of stock data

```{r warning=FALSE, message=FALSE}
#visualize
theme_set(theme_test())
TSLA %>% 
    ggplot(aes(x=Date))+
    geom_line(aes(y=Close), col='deepskyblue')+
    scale_x_date()+
    scale_x_date(date_breaks = "2 months",date_labels=("%b %y"))+
    theme(axis.text.x = element_text(
      angle = 45,
      hjust = 1
    )) +
  theme(panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()) +
  labs(
    title = "Tesla Stock Price Trend Analysis",
    x = "Date",
    y = "Stock Price (USD)"
  )

```

### Find Simple Moving Averages for 30, 100 and 200 days

```{r warning=FALSE, message=FALSE}
TSLA1 <- TSLA %>%
  mutate(SMA_30 = rollmean(Close, k = 30, fill = NA),
         SMA_100 = rollmean(Close, k = 100, fill = NA),
         SMA_200 = rollmean(Close, k = 200, fill = NA))
```

### Plot moving averages along with closing prices

```{r}
# Create breaks and labels for x-axis
tesla_data_long <- TSLA1 %>%
  pivot_longer(cols = c(Close, SMA_30, SMA_100, SMA_200), names_to = "Series", values_to = "Value")
```

```{r warning=FALSE, message=FALSE}
# Plot the reshaped data
ggplot(data = tesla_data_long, aes(x = Date, y = Value, color = Series)) +
  geom_line() +
  labs(title = "Tesla Stock Price with Moving Averages (30, 100, and 200 days)",
       x = "Date",
       y = "Stock Price (USD)",
       color = "Series") +
  scale_color_manual(values = c("Close" = "blue", 
                                 "SMA_30" = "red", 
                                 "SMA_100" = "green", 
                                 "SMA_200" = "orange"),
                     labels = c("Close", "SMA 30", "SMA 100", "SMA 200")) +
  scale_x_date(date_breaks = "2 months", date_labels =("%b %y"))+
  theme_bw()+
  theme(panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position = "bottom"
)+
  theme(axis.text.x = element_text(
    angle = 45,
    hjust = 1
  ))
```

### Calculate rolling standard deviation for 100 day and perform Volatility Analysis

```{r warning=FALSE, message=FALSE}
tesla_data <- TSLA1 %>%
  mutate(Volatility_100 = rollapply(Close, width = 100, 
                                    FUN = sd, 
                                    fill = NA))

ggplot(data = tesla_data, aes(x = Date)) +
  geom_line(aes(y = Volatility_100, 
                color = "100-Day"), 
            linetype = "solid") +
  labs(title = "Tesla Stock Volatility Analysis (100 days)",
       x = "Date",
       y = "Volatility",
       color = NULL) +
  scale_x_date(date_labels=("%b %y"), breaks="2 months")+
  theme_bw()+
  theme(axis.text.x = element_text(
    angle = 45,
    hjust = 1
  ))+
  theme(panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position="none"
)
```

### Perform Seasonal Trend Analysis by calculating monthly average prices.

```{r warning=FALSE, message=FALSE}
# Create 'Month' column from 'Date' column
tesla_data <- TSLA1 %>%
  mutate(Month = format(Date, "%Y-%m"))

# Calculate average stock price for each month
monthly_avg <- tesla_data %>%
  group_by(Month) %>%
  summarize(Avg_Price = mean(Close, na.rm = TRUE))

# Modify the 'Month' column format to display in 'Year-Month' format (YYYY-MM)
monthly_avg$Month <- as.yearmon(monthly_avg$Month)

# Plot seasonal trend using scatter plot
ggplot(monthly_avg, aes(x = Month, y = Avg_Price)) +
  geom_point(color = "blue", size = 3) +
  geom_smooth(method = "lm", se = FALSE, color = "red", linetype = "dashed") +  # Add a linear trendline
  labs(title = "Seasonal Trend of Tesla Stock Price by Month",
       x = "Month",
       y = "Average Stock Price (USD)") +
  theme_bw()+
  theme(panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
)
```
